The brightness and spatial distributions of terrestrial radio sources

### 2013-01-18 articles/Article-RFI-distribution/article-rfi-distribution.tex
Removed appendix, rewrote spherical derivation

\subsection{Spherical case} \label{sec:spherical-geometry}
If we do take into account the fact that the sources lie on a sphere instead of a plane, Eq.~\eqref{eq:cumulative-distance-function} changes. In this case, the RFI sources are distributed over the surface of a sphere. If we assume their difference in height is negligible, and by using the formula for the surface area of a hemisphere and some basic geometry, one finds that
\begin{equation}\label{eq:cumulative-spherical}
\tilde{F}_{\textrm{distance} \le r} = \rho 2\pi R^2\left(1-\cos(2\arcsin \frac{r}{2R})\right) \hspace{0.4cm}\textrm{with } r \le 2R,
\end{equation}
with $R$ the radius of the Earth and $r$ the Euclidean distance between the source and the antenna. As shown in Appendix~\ref{sec:proof-spherical-case}, Eqs.~\eqref{eq:cumulative-distance-function} and \eqref{eq:cumulative-spherical} have the same value over the range they can be evaluated. Therefore, the spherical and 2-dimensional cases lead to a distribution with equal power law. Evaluating Eq.~\eqref{eq:cumulative-spherical} is limited to $r \le 2R$, but in practice the lowest amplitude of observed sources is further constrained because of the curvature of the Earth. This causes distant sources to settle below the horizon. Nevertheless, some of those sources might still be visible due to reflection of the ionosphere and other (semi-)spherical propagation effects.

@@

\appendix
\section{Calculations with spherical geometry} \label{sec:proof-spherical-case}
As stated in \S\ref{sec:spherical-geometry}, the distance distribution of sources for a flat geometry and a spherical surface are equal for distances smaller than the diameter of the Earth. We proof this by showing that the spherical distance function $\tilde{F}_{\textrm{distance} \le r}(r)$ of Eq.~\ref{eq:cumulative-spherical} equals the flat distance function $F_{\textrm{distance} \le r}(r)$ of Eq.~\ref{eq:cumulative-distance-function}.
\begin{proof}
\begin{eqnarray} \notag
\tilde{F}_{\textrm{distance} \le r}(r) & = & \rho 2 \pi R^2 \left[1  - \cos\left(2 \arcsin \frac{r}{2R}\right)\right] \\\notag
& & \textit{\{\hspace{1mm}Using $\cos 2\phi = 1 - 2\sin^2 \phi$\hspace{1mm}\}} \\\notag
& = & \rho 2 \pi R^2 \left[1 - \left(1 - 2 \sin^2(\arcsin \frac{r}{2R})\right)\right] \\\notag
& = & \rho 2 \pi R^2 \left[2 \sin^2(\arcsin \frac{r}{2R})\right] \\\notag
& & \textit{\{\hspace{1mm}Assuming $r \le 2R$\hspace{1mm}\}} \\\notag
& = & \rho 2 \pi R^2 \left[2 \left(\frac{r}{2R}\right)^2\right] \\\notag
& = & \rho \pi r^2 \\
& = & F_{\textrm{distance} \le r}(r)
\end{eqnarray}
\end{proof}

>>>>

\subsection{Spherical case} \label{sec:spherical-geometry}
If we do take into account the fact that the sources lie on a sphere instead of a plane, Eq.~\eqref{eq:cumulative-distance-function} changes. In this case, the RFI sources are distributed over the surface of a sphere. If we assume their difference in height is negligible, one can use the formula for the surface area of a spherical cap,
\begin{equation}\label{eq:spherical-cap-area}
 A_\textrm{cap} = \pi \left( a^2 + h^2 \right) \hspace{0.4cm}\textrm{with } a \le R \wedge h \le 2R,
\end{equation}
with $R$ the radius of the Earth, $a$ the radius of the base of the cap and $h$ the height of the cap. Because the Euclidean distance $r$ between the source and the antenna is $r^2 = a^2 + h^2$, the resulting number of sources within Euclidean distance $r$ becomes
\begin{equation} \label{eq:cumulative-spherical}
\tilde{F}_{\textrm{distance} \le r}(r) = \rho A_\textrm{cap} = \rho \pi r^2 \hspace{0.4cm}\textrm{with } r \le 2R,
\end{equation}
which equals the formule for a planar source distribution given in Eq.~\eqref{eq:cumulative-distance-function}. Therefore, the spherical and two-dimensional cases lead to a distribution with equal power law. Evaluating Eq.~\eqref{eq:cumulative-spherical} is limited to $r \le 2R$, but in practice the lowest amplitude of observed sources is further constrained because of the curvature of the Earth. This causes distant sources to settle below the horizon. Nevertheless, some of those sources might still be visible due to reflection of the ionosphere and other (semi-)spherical propagation effects.

@@
